Quantum internet \& quantum machine learning

### 2019-03-05 Nana QML section/QinternetQML.tex
minor

%% Language and font encodings
\usepackage[english]{babel}
\usepackage{adjustbox}

%% Useful packages
\usepackage[colorlinks, breaklinks]{hyperref}
\usepackage{hhline}

\title{Quantum internet and quantum machine learning}

@@

In the coming quantum age we can envision three possible ways where quantum resources can come into our internet: quantum communication, quantum processing at individual nodes and data that is naturally quantum in origin. To begin, we first make a classification of the four different types of networks we can have: see Table~\eqref{tab:CommTable}. We can classify them into (i) CC (classical data and processing over classical network), or the current internet; (ii) CQ (classical data and processing over quantum network); (iii) QC (quantum data and processing over classical network) and (iv) QQ (quantum data and processing over quantum network), or a fully-quantum internet.

@@

There are 3 mains paradigms for machine learning: supervised, unsupervised and reinforcement learning. Supervised learning relies on having training data from which to form inferences and to make predictions about new incoming `test' data. On the other hand, unsupervised learning algorithms make inferences using the data at hand without access to a training stage, much like a student learning without an instructor. Reinforcement learning operates using a different framework and aims to find the best action steps in a particular environment to maximize a given reward.

@@

%\noindent $\bullet$ Two main classes of machine learning algorithms: discriminative and generative. Discriminative learning is about finding a description of the conditional probability ($P(A|B)$), i.e. given test $B$, what is the probability that it belongs to $A$? Generative modelling deals with generating new instances of the same class, so tries to recover the joint probability $P(A, B)$ if given a class set $A$. \\

@@

In particular, to identify and localise unusual behaviour in a network which can be due to natural faults or an adversarial party, network anomaly detection methods can be employed \cite{ahmed2007machine}, \cite{fraley2017promise}, \cite{joseph2013machine}. Since the results can be sensitive to the training datasets used, it is important to examine which datasets are most appropriate for one's particular applications (see a review of datasets for anomaly detection \cite{yavanoglu2017review}). In particular, there have been many proposals on using anomaly detection in the network intrusion domain. However, this approach has been criticized for its uses in real-life applications since it's often difficult to distinguish anomalies
related to intrusions from those related to other factors and the complexity of real-life networks may make it too difficult to define what is a normal signal \cite{sommer2010outside}.

@@

Machine learning algorithms themselves are vulnerable to security attacks. This comes under the field of adversarial machine learning \cite{huang2011adversarial}. There are two main types of attacks: attacks of the test data (evasion) and attacks of the training data and the machine learning models (poisoning). In real-life deployments of machine learning, data often comes from different sources which makes adversarial attacks more probable. It has been discovered that many machine learning algorithms are in fact vulnerable to adversarial attacks, the first discovered in \cite{szegedy2013intriguing}. A large proportion of the literature then focuses on the details of specific algorithms: the detection of adversaries, their different methods of attack and the particular defenses to those attacks \cite{kurakin2018adversarial}. However, recently, more foundational work has emerged to try to explain the origins of this vulnerability as arising from the high dimensionality of the data involved \cite{goodfellow2014explaining, gilmer2018adversarial,mahloujifar2018curse}.

@@

Predicting network traffic is becoming increasingly important especially in diverse and complex networks. It is commonly addressed using time-series forecasting (TSF) methods. This can make use of either statistical analysis models or supervised machine learning methods \cite{bermolen2009support, chabaa2010identification, cortez2006internet}]. Non-TSF methods also exist \cite{chen2016predicting, li2016inter}.

@@

%\noindent $\bullet$ Resource allocation using ML. 
% network traffic control systems \cite{fadlullah2017state}
\subsection{Distributed machine learning}
Distributed machine learning is the scenario where communication between computing devices are necessary to execute the machine learning algorithm of interest.

These distributed algorithms can be executed over an internet and become highly relevant in the following scenarios: (i) When data used in the training and/or testing of the machine learning algorithm originates from different sources. This is the naturally-distributed setting. In these cases, it is also possible for the loading of all data onto one machine and transferring data may be too costly and interrupts the workflow. It can even be the case that this loading time is greater than computational time, so distributed computation is more resource-efficient; (ii) When data is too large to be stored on the RAM of a single machine; (iii) When fault tolerance becomes important. For instance, for sensitive data, if data is stored in multiple places, there is less likelihood of data corruption from any single source and the data would still be available if one source fails.

@@

However, cautious usage is required. For example, here are some cases when one \textit{shouldn't} use distributed machine learning: (i) When communication and synchronisation between the distributed parties provides a bottleneck for the computation; (ii) When writing and running a distributed program is too complicated; (iii) When one can run the same algorithm on a multi-core machine. This is possible with smart data sampling, offline
schemes and efficient parallel codes.

\section{Classical data and machine learning with quantum resources}
There are at least three broad ways in which we can employ quantum resources for classical data over a network: (i) using quantum resources to enhance data processing at individual nodes (ii) using quantum resources to assist in security and/or (iii) using quantum resources to assist in communication.

@@

The first set of these algorithms have chiefly relied on assuming completely quantum devices, keeping coherence throughout the computation and could require full fault-tolerance. For those algorithms claiming up to exponential quantum speedups for supervised algorithms (see \cite{biamonte2017quantum, ciliberto2018quantum}),  the HHL algorithm \cite{harrow2009quantum} for matrix-inversion is often used. However, HHL has a number of drawbacks that make them impractical for near-term quantum devices: (i) the ability to efficiently encode the classical data into quantum states and into quantum memory \cite{aaronson2015read}; (ii) effective read-out of the final quantum state \cite{aaronson2015read}; (iii) generally requiring high circuit-depth and (iv) restrictions on the sparsity and condition numbers of the matrix.

Although later developments have tried to circumvent restrictions on sparsity and focus on low-rank matrices instead (e.g. quantum principal component analysis for low-rank matrices \cite{lloyd2014quantum}), recent work on quantum-inspired classical algorithms have demonstrated efficient classical can exist in these cases \cite{tang2018quantum, gilyen2018quantum, chia2018quantum}. In fact, classical sampling methods developed in \cite{tang2018quantum} for quantum-inspired machine learning algorithms suggest that classical methods for linear algebra problems in low-dimensions (used in machine learning for instance) are likely to find efficient classical algorithms. Although these classical sampling methods are not yet more practical than existing classical sampling methods, they are still more realistic than their quantum counterparts in (i) and (ii).

@@

Both QAOA and VQE can be considered as part of the same framework and their optimisation part (which can be considered only as a component and not the entirety of machine learning) is performed classically. One begins with an ansatz quantum state. An unitary with classically-tunable parameters is then applied to this state and an observable whose expectation value representing the cost function for the problem is subsequently measured. The classical parameters of the unitary are then iteratively tuned until one reaches the lowest value of the cost function (i.e., ground state of a given Hamiltonian), for instance using the classical gradient-descent algorithm. \

@@

\subsection{Distributed quantum machine learning}
The reasons to consider distributed quantum machine learning are similar to those for distributed classical machine learning. Suppose one wishes to perform distributed machine learning, either because the given data is naturally distributed or there is limited processing power on any given device. For this purpose, there are existing protocols for implementing general distributed quantum algorithms that could also be helpful in delegating quantum machine learning algorithms, e.g., \cite{beals2013efficient}.

@@

It is possible to process quantum data over either a classical or a quantum network. Techniques from classical machine learning methods for quantum data may assist in the communication of quantum data over a classical network while quantum learning protocols may be more appropriate over a quantum network. We are at the very beginning of this area of investigation. It is yet unclear exactly if and how these methods may be applicable and these serve as tantalizing inspiration for future study.

@@

In a future quantum internet with entangled quantum networks, it is desirable to find the optimal methods of creating the types and extent of entanglement required with respect to resource constraints. It is also desirable this process may be automated. Recently, such automated methods based on classical reinforcement learning \cite{alexey} have been proposed, to experimentally create a variety of entangled states with more efficiency. This provides an exciting beginning for automated design of future quantum internet protocols.

@@

%\textbf{Network performance optimisation}
%\section{Quantum machine learning and classical networks}
%Individual quantum machine learning algorithms with speed-ups and only classical transmission of data (e.g. classical data sent to centres which then create quantum state resources). Give a few names of algorithms with speed-ups and include names of reviews. \\

%The second part are quantum algorithms which might help with classical network problems (e.g. classical routing...etc). Also may help with cybersecurity (e.g. quantum anomaly detection).

@@

%Name if there are any advantages in using classical machine learning for quantum network problems. Mention that other classical (not necessary ML) methods used to simulate quantum networks(e.g. add SimulaQron). \\

@@

%General results in distributed quantum computation (e.g. Eisert et al; Beals et al \cite{beals2013efficient}). \\

@@

%\subsection{Quantum traffic engineering}
%\subsection{Quantum network performance optimisation} 
%\section{Outlook}
%Emergent intelligent properties of quantum networks and fully automated, self-configuring  quantum networks...
\clearpage
\begin{table}[h]
\centering
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{llll}
Data/Network & Classical & Quantum \\\hline
Classical & Current internet (CC) & Classical data in quantum network (CQ) \\
Quantum & Quantum data in classical network (QC) & Fully quantum internet (QQ)
\end{tabular}
\end{adjustbox}
\caption{\label{tab:CommTable} Classical and quantum data in a network}
\end{table}
\clearpage

\begin{table}[h]
\centering
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{lllp{5cm}}
Network  & Classical data & Quantum resources  & Quantum data \\
concerns & & and classical data & \\\hhline{|=|=|=|=|}
Individual & Classical   & Quantum-enhanced  & Quantum learning  \\ 
computing & machine learning & machine learning \\ \hline 
Security and faults & Machine learning for anomalies  & Adversarial quantum & Anomaly detection and  \\
& and faults: detection and prediction; & machine learning  & change point detection \\
& Adversarial machine learning &  & for quantum data \\ \hline 
Routing and traffic & Machine learning for \\
& traffic prediction, classification, & Open problems  & Open problems \\
& congestion control and routing\\ \hline 
Distributed  & Distributed machine learning & Distributed quantum & Distributed quantum learning \\
computing & & machine learning\\ \hline 
 Communication & Data compression & Open problems & Data compression and \\
 & and machine learning & & quantum machine learning \\ \hline 
\end{tabular}
\end{adjustbox}
\caption{\label{tab:CommTable2} Classical and quantum machine learning applications in classical and quantum networks. Almost all of the categories here are very new and open to exploration in the quantum domain.}
\end{table}

>>>>

%% Language and font encodings
\usepackage[english]{babel}

%% Useful packages
\usepackage[colorlinks, breaklinks]{hyperref}

\newcommand{\captionspacetab}[0]{\vspace{-0.6em}}

\title{Quantum internet \& quantum machine learning}

@@

In the coming quantum age we can envision three possible ways where quantum resources can come into our internet: quantum communication, quantum processing at individual nodes and data that is naturally quantum in origin. To begin, we first make a classification of the four different types of networks we can have: see Table~\eqref{tab:CommTable}. We can classify them into:

\begin{itemize}
\item CC: classical data and processing over classical network, or the current internet.
\item CQ: classical data and processing over quantum network.
\item QC: quantum data and processing over classical network.
\item QQ: quantum data and processing over quantum network, or a fully-quantum internet.
\end{itemize}

@@

There are 3 mains paradigms for machine learning: supervised, unsupervised and reinforcement learning. Supervised learning relies on having training data from which to form inferences and to make predictions about new incoming `test' data. On the other hand, unsupervised learning algorithms make inferences using the data at hand without access to a training stage, much like a student learning without an instructor. Reinforcement learning operates using a different framework and aims to find the best action steps in a particular environment to maximise a given reward.

@@

%\noindent $\bullet$ Two main classes of machine learning algorithms: discriminative and generative. Discriminative learning is about finding a description of the conditional probability ($P(A|B)$), i.e given test $B$, what is the probability that it belongs to $A$? Generative modelling deals with generating new instances of the same class, so tries to recover the joint probability $P(A, B)$ if given a class set $A$. \\

@@

In particular, to identify and localise unusual behaviour in a network which can be due to natural faults or an adversarial party, network anomaly detection methods can be employed \cite{ahmed2007machine, fraley2017promise, joseph2013machine}. Since the results can be sensitive to the training datasets used, it is important to examine which datasets are most appropriate for one's particular applications (see a review of datasets for anomaly detection \cite{yavanoglu2017review}. In particular, there have been many proposals on using anomaly detection in the network intrusion domain. However, this approach has been criticised for its uses in real-life applications since it's often difficult to distinguish anomalies
related to intrusions from those related to other factors and the complexity of real-life networks may make it too difficult to define what is a normal signal \cite{sommer2010outside}.

@@

Machine learning algorithms themselves are vulnerable to security attacks. This comes under the field of adversarial machine learning \cite{huang2011adversarial}. There are two main types of attacks: attacks of the test data (evasion) and attacks of the training data and the machine learning models (poisoning). In real-life deployments of machine learning, data often comes from different sources which makes adversarial attacks more probable. It has been discovered that many machine learning algorithms are in fact vulnerable to adversarial attacks, the first discovered in \cite{szegedy2013intriguing}. A large proportion of the literature then focuses on the details of specific algorithms: the detection of adversaries, their different methods of attack and the particular defences to those attacks \cite{kurakin2018adversarial}. However, recently, more foundational work has emerged to try to explain the origins of this vulnerability as arising from the high dimensionality of the data involved \cite{goodfellow2014explaining, gilmer2018adversarial, mahloujifar2018curse}.

@@

Predicting network traffic is becoming increasingly important especially in diverse and complex networks. It is commonly addressed using time-series forecasting (TSF) methods. This can make use of either statistical analysis models or supervised machine learning methods \cite{bermolen2009support, chabaa2010identification, cortez2006internet}. Non-TSF methods also exist \cite{chen2016predicting, li2016inter}.

@@

%\noindent $\bullet$ Resource allocation using ML. 
% network traffic control systems \cite{fadlullah2017state}

\subsection{Distributed machine learning}

Distributed machine learning is the scenario where communication between computing devices are necessary to execute the machine learning algorithm of interest.

These distributed algorithms can be executed over an internet and become highly relevant in the following scenarios:
\begin{itemize}
\item When data used in the training and/or testing of the machine learning algorithm originates from different sources. This is the naturally-distributed setting. In these cases, it is also possible for the loading of all data onto one machine and transferring data may be too costly and interrupts the workflow. It can even be the case that this loading time is greater than computational time, so distributed computation is more resource-efficient.
\item When data is too large to be stored on the RAM of a single machine.
\item When fault tolerance becomes important. For instance, for sensitive data, if data is stored in multiple places, there is less likelihood of data corruption from any single source and the data would still be available if one source fails.
\end{itemize}

@@

However, cautious usage is required. For example, here are some cases when one \textit{shouldn't} use distributed machine learning:
\begin{itemize}
\item When communication and synchronisation between the distributed parties provides a bottleneck for the computation.
\item When writing and running a distributed program is too complicated.
\item When one can run the same algorithm on a multi-core machine. This is possible with smart data sampling, offline schemes and efficient parallel codes. 
\end{itemize}

\section{Classical data and machine learning with quantum resources}

There are at least three broad ways in which we can employ quantum resources for classical data over a network:
\begin{itemize}
\item Using quantum resources to enhance data processing at individual nodes.
\item Using quantum resources to assist in security.
\item Using quantum resources to assist in communication.
\end{itemize}

@@

The first set of these algorithms have chiefly relied on assuming completely quantum devices, keeping coherence throughout the computation and could require full fault-tolerance. For those algorithms claiming up to exponential quantum speedups for supervised algorithms (see \cite{biamonte2017quantum, ciliberto2018quantum}),  the HHL algorithm \cite{harrow2009quantum} for matrix-inversion is often used. However, HHL has a number of drawbacks that make them impractical for near-term quantum devices:
\begin{itemize}
\item The ability to efficiently encode the classical data into quantum states and into quantum memory \cite{aaronson2015read}.
\item Effective read-out of the final quantum state \cite{aaronson2015read}.
\item Generally requiring high circuit-depth.
\item Restrictions on the sparsity and condition numbers of the matrix.
\end{itemize}

Although later developments have tried to circumvent restrictions on sparsity and focus on low-rank matrices instead (e.g quantum principal component analysis for low-rank matrices \cite{lloyd2014quantum}), recent work on quantum-inspired classical algorithms have demonstrated efficient classical can exist in these cases \cite{tang2018quantum, gilyen2018quantum, chia2018quantum}. In fact, classical sampling methods developed in \cite{tang2018quantum} for quantum-inspired machine learning algorithms suggest that classical methods for linear algebra problems in low-dimensions (used in machine learning for instance) are likely to find efficient classical algorithms. Although these classical sampling methods are not yet more practical than existing classical sampling methods, they are still more realistic than their quantum counterparts in (i) and (ii).

@@

Both QAOA and VQE can be considered as part of the same framework and their optimisation part (which can be considered only as a component and not the entirety of machine learning) is performed classically. One begins with an ansatz quantum state. A unitary with classically-tuneable parameters is then applied to this state and an observable whose expectation value representing the cost function for the problem is subsequently measured. The classical parameters of the unitary are then iteratively tuned until one reaches the lowest value of the cost function (i.e ground state of a given Hamiltonian), for instance using the classical gradient-descent algorithm. \

@@

\subsection{Distributed quantum machine learning}
The reasons to consider distributed quantum machine learning are similar to those for distributed classical machine learning. Suppose one wishes to perform distributed machine learning, either because the given data is naturally distributed or there is limited processing power on any given device. For this purpose, there are existing protocols for implementing general distributed quantum algorithms that could also be helpful in delegating quantum machine learning algorithms, e.g \cite{beals2013efficient}.

@@

It is possible to process quantum data over either a classical or a quantum network. Techniques from classical machine learning methods for quantum data may assist in the communication of quantum data over a classical network while quantum learning protocols may be more appropriate over a quantum network. We are at the very beginning of this area of investigation. It is yet unclear exactly if and how these methods may be applicable and these serve as tantalising inspiration for future study.

@@

In a future quantum internet with entangled quantum networks, it is desirable to find the optimal methods of creating the types and extent of entanglement required with respect to resource constraints. It is also desirable this process may be automated. Recently, such automated methods based on classical reinforcement learning \cite{alexey} have been proposed to experimentally create a variety of entangled states with more efficiency. This provides an exciting beginning for automated design of future quantum internet protocols.

@@

%\textbf{Network performance optimisation}
%\section{Quantum machine learning and classical networks}
%Individual quantum machine learning algorithms with speed-ups and only classical transmission of data (e.g classical data sent to centres which then create quantum state resources). Give a few names of algorithms with speed-ups and include names of reviews. \\

%The second part are quantum algorithms which might help with classical network problems (e.g classical routing...etc). Also may help with cybersecurity (e.g quantum anomaly detection).

@@

%Name if there are any advantages in using classical machine learning for quantum network problems. Mention that other classical (not necessary ML) methods used to simulate quantum networks(e.g add SimulaQron). \\

@@

%General results in distributed quantum computation (e.g Eisert et al; Beals et al \cite{beals2013efficient}). \\

@@

%\subsection{Quantum traffic engineering}
%\subsection{Quantum network performance optimisation} 
%\section{Outlook}
%Emergent intelligent properties of quantum networks and fully automated, self-configuring  quantum networks...

\begin{table*}[htbp!]
\begin{tabular}{|l|l|l|l|}
\hline
Data/Network & Classical & Quantum \\
\hline
\hline
Classical & Current internet (CC) & Classical data in quantum network (CQ) \\
Quantum & Quantum data in classical network (QC) & Fully quantum internet (QQ)\\
\hline
\end{tabular}
\captionspacetab \caption{\label{tab:CommTable} Classical and quantum data in a network}
\end{table*}

\begin{table*}[htbp!]
\begin{tabular}{|l|l|l|l|}
\hline
Network  & Classical data & Quantum resources  & Quantum data \\
concerns & & and classical data & \\
\hline
\hline
Individual & Classical   & Quantum-enhanced  & Quantum learning  \\ 
computing & machine learning & machine learning & \\
\hline 
Security and faults & Machine learning for anomalies & Adversarial quantum & Anomaly detection and  \\
& and faults: detection and prediction; & machine learning  & change point detection \\
& Adversarial machine learning &  & for quantum data \\
\hline 
Routing and traffic & Machine learning for  & Open problems & Open problems \\
& traffic prediction, classification, &   &  \\
& congestion control and routing & & \\
\hline 
Distributed  & Distributed machine learning & Distributed quantum & Distributed quantum learning \\
computing & & machine learning & \\
\hline 
 Communication & Data compression & Open problems & Data compression and \\
 & and machine learning & & quantum machine learning \\
\hline 
\end{tabular}
\captionspacetab \caption{\label{tab:CommTable2} Classical and quantum machine learning applications in classical and quantum networks. Almost all of the categories here are very new and open to exploration in the quantum domain.}
\end{table*}
